Demographic history and patterns of selection in maize during and since domestication

### 2015-07-30 manuscript.tex
even closer

\title{The impacts of demography on linked selection in maize}
\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
    California, Davis, CA, USA}, Li Wang\affil{2}{Iowa State University, Ames, IA, USA}, Arun
  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center and Center for population biology, University of
    California, Davis, CA, USA} }

@@

\begin{abstract}
  Unique selective and demographic forces operate on plants during domestication. These forces interact to generate the patterns of DNA variability that are persistent in crop species today. To quantify this interplay, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide. We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.  We obtained a complete estimate of the population size fluctuation and other demographic parameters experienced by maize as it was domesticated from teosinte. Here, we show that maize experienced a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication. We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize and teosinte apart. Also, we find that reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize. However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte. Our results demonstrate that in domesticated species, demography and selection are constantly fluctuating through time.
\end{abstract}

\dropcap{D}omesticated plant species evolve in a unique fashion
compared to their wild counterparts \textcolor{red}{(ref here)}. This
is a result of both the anthropomorphic nature of artificial selection on
domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
bottleneck(s) that they tend to have experienced
\cite{ross2007}. However, the
complex interplay between selective pressures and demographic
limitations, and the impact that this interplay has on identifying
selection and understanding demography, is not fully understood. Although a large body of
research that involves searching the genomes of domesticated species for evidence
of positive selection exists \cite{hufford2012, he2011, OTHERS?}, these studies tend to focus on
identifying or mapping particular genes or regions that play an
important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
selection have on whole-genome patterns of genetic variability therefore remains limited.

For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
parameters of a population such as effective population size, structure, and expansion
history \cite{luikart2003, gutenkunst2009}. However, researchers have
called into question whether or not there are reasonable approaches
for identifying neutral regions of the genome, since the effects of
selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
\emph{Drosophila} it appears that the majority of the genome is
impacted by the effects of selection through linkage
\cite{sella2009}. This observation begs the question, how can
demographic parameters be estimated independently of
selective parameters? Human researchers have attempted to address this
problem by limiting analyses to only sites far from genes
\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
difficult to be certain that sites even in gene-poor regions of the
genome are not influenced by linked selection. Ultimately, an understanding
of which sites have the potential to be adaptive, and how these affect
genome-wide patterns of variability through linkage, is required for
reasonable demographic inference.

Maize represents an excellent organism to study these
phenomena. Maize is a species of tremendous importance worldwide as
both a staple crop \cite{SOMETHING} and as a model for
understanding crop evolution \cite{SOMETHINGS}. Broadly speaking, archaeological and genetic studies have
established that maize domestication is likely to have taken place in
Central Mexico approximately 9,000 years bp
\cite{SOMETHING,SOMETHING}. Teosinte, the most
recent wild ancestor to maize, remains extant throughout much of the
Americas \cite{SOMETHING}. Additionally, several large-effect
domestication loci \cite{SOMETHINGS} and putative domestication
regions \cite{hufford2012} have been identified. But despite all that is
known about maize domestication, the parameters of the
domestication process remain uncertain. Specifically, the size of the
maize domestication bottleneck has not been estimated independently of
the bottleneck's duration, nor are there sequence-based estimates of the effective
population size of modern maize. Sequence information from maize and
teosinte plants may therefore be utilized to address these questions.

@@

\section{Results}
\subsection{Patterns of diversity differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \textcolor{red}{(several references
  here)}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.0115
(p$\approx$0). These observations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection is likely
reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

@@

\subsection{Demography of maize domestication}
To explore whether differences in purifying selection between maize
and teosinte can be explained by demographic processes, we
first estimated the parameters of maize domestication using large-scale
sequencing data involving 23 inbred maize
landraces and 13 teosinte inbred lines included in the HapMap 2 panel
\cite{chia2012} \textcolor{red}{(supplemental table here)}. The maize
lines were collected from across the Americas and the teosinte lines
came from central Mexico. Before estimating demography, we compared
the site frequency spectrum (SFS) in genic and non-genic regions, and
observed substantial differences in the evolution of these classes of
sites. For both maize and teosinte, the SFS within genes showed a
dearth of low-frequency alleles and Tajima's D
\textcolor{red}{(reference here)} was therefore shifted
to more positive values \textcolor{red}{(figure here)}. This is consistent with the aforementioned purifying
selection and indicates that genic regions are not evolving
neutrally. This observation is consistent with previous research \cite{hufford2012}. Therefore, for demographic modeling we restricted analysis
to sites at least 5kb from gene start/stop postions.

@@

The most likely model suggested an ancestral for $\theta$ ($4N\mu$) of
0.014734, which is similar to previous estimates
\textcolor{red}{(references here)}. Assuming a mutation rate of $\mu =
3 \times 10^{-8}$ \textcolor{red}{(reference here)}, this suggests an effective
population size of $N_a = 122,783$ individuals, a population split
$T_B = 15,523$ generations in the past, a maize bottleneck
effective populations size of $N_b = 6,455$ individuals (5.26\% of
$N_a$), a modern maize effective population size of $N_m = 366,973$
individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
maize, and $M_{mt} = 1.72$ migrants per generation from maize to
teosinte (Figure \ref{fig:bottleneck}). We note that a population
split over 15 thousand generations before present precedes estimates
from
archaeological data which suggests maize domestication began
approximately 9,000 years before present \textcolor{red}{(reference
  here)}. This could reflect teosinte population structure that was present before
domestication. Also, note that the genetic time of the population
split must precede morphological changes that could be identified
morphologically.   Additionally, because recent expansion is
most evidenced by rare alleles, and since these data provide low
power to detect rare alleles, we expect that the estimate of  $N_m =
366,973$, or $\sim 3N_a$, is likely an underestimate.

@@

\subsection{Demography of domestication}
Since we observed that classical hard sweeps do not explain patterns of maize diversity, it is perhaps unsurprising that we instead found evidence of demographic history contributing substantially to genetic diversity. Our estimate of the demographic history of maize, which is the first estimate of maize demographic history to utilize whole-genome-sequence (WGS) data, allowed us to disentangle the confounding parameters of bottleneck strength and bottleneck intensity. This confounding limited previous domestication estimates \cite{wright2005}. By utilizing the two-dimensional site frequency spectrum (SFS) between maize and teosinte for inference \cite{gutenkunst2009} we estimate that the maize effective population size during the bottleneck was approximately 5\% that of teosinte. Elevated drift during a bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, as is predicted during a domestication bottleneck \cite{somebody}.

After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte. Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \cite{somebody} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte. From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck. A back-of-the-envelope calculation of modern $N_e$ can be made by assuming there are 47.9 million ha of landrace maize in production \cite{cimmyt1999}, 42,000 plants per ha for open pollinated maize varieties \cite{van2010}, and conservatively that no more than one in 1,000 plants contribute gametes to the next generation. This calculation implies a modern $N_e \geq  2$ billion individuals, and incorporates no individuals from hybrid breeding programs. If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domesticaiton expansion in maize $N_e$ is enormous. Patterns of singleton diversity in maize, as discussed in more detail in the next section, demonstrate that this recent expansion has a notable impact on the maize genome.

\subsection{Demography strongly impacts efficiency of purifying selection}
The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint. Theory demonstrates that purifying selection is more efficient in a large population than in a small one \cite{somebody}, and this observation most likely reflects that prediction. More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past. Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively in maize as in teosinte.

@@

A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte. This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{something}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds. These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected. The sample we studied contained only open-pollinated varieties of both maize and teosinte, so we cannot explicitly test this prediction.

@@

\subsection{Demographic inference}
We used dadi \cite{gutenkunst2009} to estimate the parameters of maize
domestication based on diffusion approximation of the 2-dimensional
SFS for maize and teosinte. Our first step was to specify a model, as
shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
population with effective population size $N_a$ individuals. The
ancestral population split into
two distinct populations, one of maize and the other of teosinte,
$T_b$ generations in the past. The maize population immediately shrank
to size $N_b$, and then grew exponentially to size $N_{maize}$
today. From the split until today, the teosinte effective population
size remained constant. Each year, $M_{mt}$ individuals migrated from
maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm} $. Python code specifying this model is
provided in \textcolor{red}{(SUPPLEMENT)}. We implimented 1,000 dadi
optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
optimizer. In dadi optimizations,
initial values for $N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
bounds were set at 1e-7, 1e-7,  0, 1e-10, and
1e-10, respecively, while upper bounds were respectively set at 2,
200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
lower bounds because it is estimated from equilibrium ancestral theta
($\theta = 4N_a\mu$), which dadi determines automatically regardless
of model. Population sizes were converted from dadi's
units using a mutation rate estimate of 3e-8
\textcolor{red}{(REFERENCE)}. Among the 1,000 dadi runs, parameters
from that with the lowest log-likelihood were taken as estimates of
the parameters of the domestication process.

>>>>

\title{Demographic history and patterns of selection in maize during and since domestication}
\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
    California, Davis, CA, USA}, Li Wang\affil{2}{Iowa State University, Ames, IA, USA}, Arun
  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center and Center for population biology, University of
    California, Davis, CA, USA} }

@@

\begin{abstract}
  Unique selective and demographic forces operate on plants during domestication. These forces interact to generate the patterns of DNA variability that are persistent in crop species today. To quantify this interplay, we investigated genetic diversity in maize, one of the most important crops for food, feed, and fuel world-wide. We utilized whole genome sequence data from 23 maize and 13 teosinte individuals to make inferences.  We obtained a complete estimate of the population size fluctuations and other demographic parameters experienced by maize as it was domesticated from teosinte. Here, we show that maize experienced a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid population size expansion post-domestication. We observe that hard sweeps, specifically positive selection on new genic mutations, are not the primary force driving maize evolution. Also, we find that a reduced population size during domestication decreased the efficiency of purifying selection at purging deleterious alleles from maize. However, expansion after domestication has since increased the efficiency of purifying selection to levels superior to those seen in teosinte. Our results demonstrate that in domesticated species, demography and selection are constantly fluctuating through time.
\end{abstract}

\dropcap{D}omesticated plant species evolve in a unique fashion
compared to their wild counterparts \cite{doebley2006}. This
is a result of both the anthropomorphic nature of artificial selection on
domesticates \cite{purugganan2009} as well as the demographic characteristics of the domestication
bottleneck(s) that they tend to have experienced
\cite{ross2007}. However, the
complex interplay between selective pressures and demographic
limitations, and the impact that this interplay has on identifying
selection and understanding demography, is not fully understood. Although a large body of
research that involves searching the genomes of domesticated species for evidence
of positive selection exists \cite{hufford2012, he2011, vigouroux2002, chapman2008}, these studies tend to focus on
identifying or mapping particular genes or regions that play an
important role in phenotypic evolution. In contrast, knowledge regarding the impacts that demography and
selection have on whole-genome patterns of genetic variability therefore remains limited.

For instance, supposed neutrally-evolving DNA is often used to estimate historical demographic
parameters of a population such as effective population size, structure, and expansion
history \cite{luikart2003, gutenkunst2009}. However, researchers have
called into question whether or not there are reasonable approaches
for identifying neutral regions of the genome, since the effects of
selection can be wide-ranging \cite{li2012, slotte2014}. For example, in
\emph{Drosophila} it appears that the majority of the genome is
impacted by the effects of selection through linkage
\cite{sella2009}. A natural next question, therefore, is how can
demographic parameters be estimated independently of
selective parameters? Human researchers have attempted to address this
problem by limiting analyses to only sites far from genes
\cite{gazave2014}, but as \emph{Drosophila} demonstrates, it is
difficult to be certain that sites even in gene-poor regions of the
genome are not influenced by linked selection. Ultimately, an understanding
of which sites have the potential to be adaptive, and how these affect
genome-wide patterns of variability through linkage, is required for
reasonable demographic inference.

Maize represents an excellent organism to study these
phenomena. Maize is a species of tremendous importance worldwide as
both a staple crop \cite{shiferaw2011} and as a model for
understanding plant evolution \cite{strable2009}. Broadly speaking, archaeological and genetic studies have
established that maize domestication is likely to have taken place in
 Mexico approximately 9,000 years bp
\cite{smith1995,matsuoka2002}. Teosinte, the most
recent wild ancestor to maize, remains extant throughout much of the
Americas \cite{wilkes1967}. Additionally, several large-effect
domestication loci \cite{doebley1995, wills2013, wang2015} and putative domestication
regions \cite{hufford2012} have been identified. But despite all that is
known about maize domestication, the parameters of the
domestication process remain uncertain. Specifically, the size of the
maize domestication bottleneck has not been estimated independently of
the bottleneck's duration, nor are there sequence-based estimates of the effective
population size of modern maize. Sequence information from maize and
teosinte plants may therefore be utilized to address these questions.

@@

\section{Results}
\subsection{Patterns of diversity differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \cite{wright2005,wang1999,eyre1998}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.0115
(p$\approx$0). These observations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection is likely
reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

@@

\subsection{Demography of maize domestication}
To explore whether differences in purifying selection between maize
and teosinte can be explained by demographic processes, we
first estimated the parameters of maize domestication using large-scale
sequencing data involving 23 inbred maize
landraces and 13 teosinte inbred lines included in the HapMap 2 panel
\cite{chia2012} \textcolor{red}{(supplemental table here)}. The maize
lines were collected from across the Americas and the teosinte lines
came from central Mexico. Before estimating demography, we compared
the site frequency spectrum (SFS) in genic and non-genic regions, and
observed substantial differences in the evolution of these classes of
sites. For both maize and teosinte, the SFS within genes showed a
dearth of low-frequency alleles and Tajima's D
\cite{tajima1989} was therefore shifted
to more positive values \ref{fig:sffs}. This is consistent with the aforementioned purifying
selection and indicates that genic regions are not evolving
neutrally. This observation is consistent with previous research \cite{hufford2012}. Therefore, for demographic modeling we restricted analysis
to sites at least 5kb from gene start/stop postions.

\begin{figure}[b]
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/SFS_and_Tajima.png}
\caption{The folded SFS and Tajimas D were calculated both inside and outside of genes for maize and teosinte. In both taxa, a dearth of rare alleles was observed within genes relative to outside of genes, which led to higher values of Tajima's D outside of genes than inside.{\bf A:} Folded SFS for maize; {\bf B:} Folded SFS for teosinte; {\bf C:} Maize Tajima's D; {\bf D:} Teosinte Tajima's D.  \label{fig:sfs} }
\end{figure}

@@

The most likely model suggested an ancestral for $\theta$ ($4N\mu$) per base pair of
0.014734, which approximately matches a previous independent estimate \cite{eyre1998}. Assuming a mutation rate of $\mu =
3 \times 10^{-8}$ \cite{clark2005}, this suggests an effective
population size of $N_a = 122,783$ individuals, a population split
$T_B = 15,523$ generations in the past, a maize bottleneck
effective populations size of $N_b = 6,455$ individuals (5.26\% of
$N_a$), a modern maize effective population size of $N_m = 366,973$
individuals, $M_{tm} = 1.35$ migrants per generation from teosinte to
maize, and $M_{mt} = 1.72$ migrants per generation from maize to
teosinte (Figure \ref{fig:bottleneck}). We note that a population
split over 15 thousand generations before present precedes estimates
from
archaeological data which suggests maize domestication began
no more than 10,000 years before present \cite{smith1995}. This could reflect teosinte population structure that was present before
domestication. Also, note that the genetic time of the population
split must precede morphological changes that could be identified
morphologically.   Additionally, because recent expansion is
most evidenced by rare alleles, and since these data provide low
power to detect rare alleles, we expect that the estimate of  $N_m =
366,973$, or $\sim 3N_a$, is likely an underestimate.

@@

\subsection{Demography of domestication}
Since we observed that classical hard sweeps do not explain patterns of maize diversity, it is perhaps unsurprising that we instead found evidence of demographic history contributing substantially to genetic diversity. Our estimate of the demographic history of maize, which is the first estimate of maize demographic history to utilize whole-genome-sequence (WGS) data, allowed us to disentangle the confounding parameters of bottleneck strength and bottleneck intensity. This confounding limited previous domestication estimates \cite{wright2005}. By utilizing the two-dimensional site frequency spectrum (SFS) between maize and teosinte for inference \cite{gutenkunst2009} we estimate that the maize effective population size during the bottleneck was approximately 5\% that of teosinte. A bottleneck of this magnitude may have enabled moderately deleterious alleles to reach intermediate frequency in maize, while also increasing the probability that strongly deleterious alleles segregating at low frequency in teosinte were purged from the maize gene pool, due to elevated drift as is predicted during a domestication bottleneck \cite{tajima1989}.

After the initial bottleneck, our analysis suggests that maize $N_e$ expanded to reach levels much greater than that of teosinte. Our WGS-based estimate shows that the $N_e$ of landrace maize is at least 3X that of ancestral teosinte, while our singleton-based estimate from the the SeeDs project data \textcolor{red}{(citation)} implies an $N_e$ of close to 1 million individuals, or $\sim$8X that of ancestral teosinte. From the perspective of a new mutation, both of these estimates are biased downward, since they represent past reductions in diversity due to the aforementioned bottleneck. A back-of-the-envelope calculation of modern $N_e$ can be made by assuming there are 47.9 million ha of landrace maize in production \cite{cimmyt1999}, 42,000 plants per ha for open pollinated maize varieties \cite{van2010}, and conservatively that no more than one in 1,000 plants contribute gametes to the next generation. This calculation implies a modern $N_e \geq  2$ billion individuals, and incorporates no individuals from hybrid breeding programs. If for some reason this back-of-the-envelope calculation is an overestimate, even by several orders of magnitude, it is clear that the post-domesticaiton expansion in maize $N_e$ is enormous. Patterns of singleton diversity in maize, as discussed in more detail in the next section, demonstrate that this recent expansion has a notable impact on the maize genome.

\subsection{Demography strongly impacts efficiency of purifying selection}
The observation that maize pairwise diversity is less impacted by the distance to the nearest gene than is teosinte pairwise diversity is reasonable from a long-term evolutionary standpoint. Theory has established that purifying selection is more efficient in a large population than in a small one \cite{kimura1984}, and this observation most likely reflects that prediction. More specifically, if teosinte $N_e$ remained relatively constant while maize bottlenecked and recovered exponentially, this provides that the average $N_e$ of maize over the previous several thousand generations is much smaller than that of teosinte, regardless of how much maize has ballooned in the recent past. Therefore, our observation shows that purifying selection in maize has not purged deleterious alleles, or the neutral alleles they are linked to, as effectively in maize as in teosinte.

@@

A consequence of the inefficient purifying selection that maize experienced during its bottleneck is likely that it harbors more deleterious alleles segregating at intermediate frequency than does teosinte. This could be a part of the explanation of why maize inbreds have continued to improve over the past several decades \cite{meghji1984}; if deleterious alleles tend to be recessive and are particularly frequent, they will have increased opportunities to display their phenotypes in inbreds. These results also demonstrate that recent purifying selection in maize has become much more effective, potentially explaining the ongoing improvement of these inbreds as maize lines are continuously selected. The sample we studied contained only open-pollinated varieties of both maize and teosinte, so we cannot explicitly test this prediction.

@@

\subsection{Demographic inference}
We used dadi \cite{gutenkunst2009} to estimate the parameters of maize
domestication based on diffusion approximation of the 2-dimensional
SFS for maize and teosinte. Our first step was to specify a model, as
shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
population with effective population size $N_a$ individuals. The
ancestral population split into
two distinct populations, one of maize and the other of teosinte,
$T_b$ generations in the past. The maize population immediately shrank
to size $N_b$, and then grew exponentially to size $N_{maize}$
today. From the split until today, the teosinte effective population
size remained constant. Each year, $M_{mt}$ individuals migrated from
maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm} $. Python code specifying this model is
provided in \textcolor{red}{(SUPPLEMENT)}. We implimented 1,000 dadi
optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
optimizer. In dadi optimizations,
initial values for $N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
bounds were set at 1e-7, 1e-7,  0, 1e-10, and
1e-10, respecively, while upper bounds were respectively set at 2,
200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
lower bounds because it is estimated from equilibrium ancestral theta
($\theta = 4N_a\mu$), which dadi determines automatically regardless
of model. Population sizes were converted from dadi's
units using a mutation rate estimate of 3e-8
\cite{clark2005}. Among the 1,000 dadi runs, parameters
from that with the lowest log-likelihood were taken as estimates of
the parameters of the domestication process.
